niche

### 2020-11-26 analyses/old_scripts/niche_v2.Rmd
did so many things

### Get pariwise niche overlap
```{r eval=FALSE, include=FALSE}
### Species-level by transect
niche_overlap.sp <- net %>%
  group_by(bb.sp, plant.sp, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = plant.sp, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

@@

niche_overlap.sp.mean <- niche_overlap.sp %>%
  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

niche_overlap.sp.mean2x <- niche_overlap.sp %>%
  group_by(site, year, date, elev.mean) %>% # Average across species pairs within transects
  summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
  group_by(site, year, elev.mean) %>% # Average across transects
  summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

### Genus-level by transect
niche_overlap.gen <- net %>%
  group_by(bb.sp, plant.genus, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = plant.genus, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

@@

niche_overlap.gen.mean <- niche_overlap.gen %>%
  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

### K-type by transect
niche_overlap.ktype <- net %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, k.type.s, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  pivot_wider(names_from = k.type.s, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

niche_overlap.ktype.mean <- niche_overlap.ktype %>%
  group_by(sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)
```

### Niche-overlap by year  
```{r}
ggplot(niche_overlap.sp, aes(year, niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen, aes(year, niche.overlap)) +
  geom_boxplot()

@@

ggplot(niche_overlap.sp.mean, aes(year, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen.mean, aes(year, mean.niche.overlap)) +
  geom_boxplot()

@@

ggplot(niche_overlap.sp.mean2x, aes(year, mean.niche.overlap)) +
  geom_boxplot()

ggplot(niche_overlap.gen.mean2x, aes(year, mean.niche.overlap)) +
  geom_boxplot()

@@

### Niche-overlap by subgenus
```{r}
ggplot(niche_overlap.sp, aes(subgenus.diff, niche.overlap)) +
  geom_boxplot()

@@

ggplot(niche_overlap.sp.mean, aes(subgenus.diff, mean.niche.overlap)) +
  geom_boxplot()

@@

### Niche-overlap by tongue length differences
```{r}
ggplot(niche_overlap.sp, aes(sqrt(pbl.diff), niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

@@

ggplot(niche_overlap.sp.mean, aes(sqrt(pbl.diff), mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

@@

### Niche-overlap by elevation  
```{r}
ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.gen, aes(elev.mean, niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.ktype, aes(elev.mean, niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(elev.mean, mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean2x, aes(elev.mean, mean.niche.overlap)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~year)

@@

### Niche overlap BB/FL covars
```{r}
# Species level
ggplot(niche_overlap.sp, aes(fl.rich, distance)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

@@

### Niche overlap by time
```{r}
ggplot(niche_overlap.sp, aes(yday(date), distance)) +
  geom_point() +
  geom_smooth() +
  #facet_wrap(~sp.pair) +
  ylim(c(0,1))

@@

### Other stuff
```{r}

net_ord_genus_dist <- vegdist(net_ord_genus, method = "horn")
net_ord_genus_clust <- hclust(net_ord_genus_dist, "ward.D")
plot(net_ord_genus_clust)
groups <- cutree(net_ord_genus_clust, 4)

>>>>

### Get pariwise niche overlap
```{r eval=FALSE, include=FALSE}
### Species-level by transect
niche_overlap.sp <- net %>%
  filter(!is.na(pollen)) %>%
  group_by(bb.sp, pollen, plant.sp, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, pollen, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  #unite(ID, c(bb.sp, pollen), sep = "_"),
  pivot_wider(names_from = plant.sp, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(pollen, site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(pollen, site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

@@

niche_overlap.sp.mean <- niche_overlap.sp %>%
  group_by(pollen, sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

niche_overlap.sp.mean2x <- niche_overlap.sp %>%
  group_by(pollen, site, year, date, elev.mean) %>% # Average across species pairs within transects
  summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
  group_by(pollen, site, year, elev.mean) %>% # Average across transects
  summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

### K-type by transect
niche_overlap.ktype <- net %>%
  filter(!is.na(pollen)) %>%
  left_join(fl_traits) %>%
  group_by(bb.sp, pollen, k.type.s, site, date, year) %>%
  summarize(freq = n()) %>% # get interaction frequency for each BB-FL pair
  group_by(bb.sp, pollen, site, date, year) %>%
  mutate(prop = freq/sum(freq)) %>% # get proportional interaction frequency for each BB-FL pair
  select(-freq) %>% # drop freq variable
  #unite(ID, c(bb.sp, pollen), sep = "_"),
  pivot_wider(names_from = k.type.s, values_from = prop) %>% # spread into the kind of array vegan likes
  replace(is.na(.), 0) %>% # replace NAs with zeros
  group_by(pollen, site, date, year) %>%
  nest() %>% # nest into list-column
  mutate(data2 = map(data, function(x) column_to_rownames(x, var = "bb.sp")), # column to rownames
         dist = map(data2, function(x) vegdist(x, method = "horn")), # convert to Horn's distance matrix
         tidy_dist = map(dist, tidy)) %>% # convert to tidy array of Horn's distances
  select(pollen, site, date, year, tidy_dist) %>% # drop intermediate list-columns
  unnest(cols = c(tidy_dist)) %>% # unnest back into a single data frame
  left_join(site_data) %>% # add site data
  select(site, date, year, pollen, bb.sp1 = item1, bb.sp2 = item2, distance, elev.mean) %>% # select and rename vars
  left_join(bb_traits, by = c("bb.sp1" = "bb.sp")) %>% # add traits for bb.sp1
  rename(subgenus1 = subgenus, pbl.w1 = pbl.w, pbl.w.class1 = pbl.w.class) %>%
  left_join(bb_traits, by = c("bb.sp2" = "bb.sp")) %>% # add traits for bb.sp2
  rename(subgenus2 = subgenus, pbl.w2 = pbl.w, pbl.w.class2 = pbl.w.class) %>%
  select(site, date, year, pollen, bb.sp1, bb.sp2, subgenus1, subgenus2, distance, 
         pbl.w1, pbl.w2, pbl.w.class1, pbl.w.class2, elev.mean) %>%
  mutate(pbl.diff = abs(pbl.w1 - pbl.w2), # tongue length differences
         yday = yday(date), # yday variable will come in handy
         subgenus.diff = if_else(subgenus1 == subgenus2, "same", "different"), # binary: same or different subgenus

@@

niche_overlap.ktype.mean <- niche_overlap.ktype %>%
  group_by(pollen, sp.pair, site, year, elev.mean, pbl.diff, subgenus.diff, 
           bss, pasc, wurf, prat, soro, hort, jone, psit, mont, hypn, 
           lapi, muci, gers, mend, pyre, humi) %>%
  summarize(mean.niche.overlap = mean(niche.overlap)) %>%
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)

niche_overlap.ktype.mean2x <- niche_overlap.ktype %>%
  group_by(pollen, site, year, date, elev.mean) %>% # Average across species pairs within transects
  summarize(mean.niche.overlap = mean(niche.overlap)) %>% 
  group_by(pollen, site, year, elev.mean) %>% # Average across transects
  summarize(mean.niche.overlap = mean(mean.niche.overlap)) %>% 
  left_join(BB_covars.mean) %>%
  left_join(FL_covars.mean)
```

### Niche-overlap by year  
```{r}
ggplot(niche_overlap.sp, aes(year, niche.overlap, color = pollen)) +
  geom_boxplot()

@@

ggplot(niche_overlap.sp.mean, aes(year, mean.niche.overlap, color = pollen)) +
  geom_boxplot()

@@

ggplot(niche_overlap.sp.mean2x, aes(year, mean.niche.overlap, color = pollen)) +
  geom_boxplot()

@@

### Niche-overlap by subgenus
```{r}
ggplot(niche_overlap.sp, aes(subgenus.diff, niche.overlap, color = pollen)) +
  geom_boxplot()

@@

ggplot(niche_overlap.sp.mean, aes(subgenus.diff, mean.niche.overlap, color = pollen)) +
  geom_boxplot()

@@

### Niche-overlap by tongue length differences
```{r}
ggplot(niche_overlap.sp, aes(sqrt(pbl.diff), niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

@@

ggplot(niche_overlap.sp.mean, aes(sqrt(pbl.diff), mean.niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4))

@@

### Niche-overlap by elevation  
```{r}
ggplot(niche_overlap.sp, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.ktype, aes(elev.mean, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean, aes(elev.mean, mean.niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.ktype.mean, aes(elev.mean, mean.niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

ggplot(niche_overlap.sp.mean2x, aes(elev.mean, mean.niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~year)

ggplot(niche_overlap.ktype.mean2x, aes(elev.mean, mean.niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  facet_wrap(~year)

@@

### Niche overlap BB/FL covars
```{r}
# Species level
ggplot(niche_overlap.sp, aes(fl.rich, niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 4)) +
  #facet_wrap(~sp.pair) +
  #facet_wrap(~year) +
  ylim(c(0,1))

@@

### Niche overlap by time
```{r}
ggplot(niche_overlap.sp, aes(yday(date), niche.overlap, color = pollen)) +
  geom_point() +
  geom_smooth() +
  #facet_wrap(~sp.pair) +
  ylim(c(0,1))

@@

### Other stuff
```{r}
net_ord_genus_dist <- vegdist(net_ord_genus, method = "horn")
net_ord_genus_clust <- hclust(net_ord_genus_dist, "ward.D")
plot(net_ord_genus_clust)
groups <- cutree(net_ord_genus_clust, 4)
